A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-03-29 casm.tex
Introduction updated, and new references to bib file. Included work of Li and Cao that has similar conlcusions to our work.

The analysis of cooperation in scientific research has been the subject of a
number of studies
\cite{VermeulenPenders:2013,Newman:2001,Newman:2004,Elango-etal:2012,HernandezLemus:2013,Strasser:2006,Strasser:2012}. This
is not surprising since cooperation and competition are quite important in
today's academic success. Moreover, the quantitative analysis of cooperation in
science also has a long history of more than 50 years
\cite{Mali-etal:2012,Scharnhorst-etal:2012}. Particularly appealing to researchers studying
this topic has been the use of network analysis, especially in the area of
bibliometrics and scientometrics. The approach we follow here articulates game theory
together with complex network analysis. Putting together game-theoretic modeling
on complex networks is common in studies about the emergence of cooperation
\cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}.
However, in the context of science studies, it is far from being a unified, active
field of research. There is no common set of questions and problems or a
coherent body of models and theory aimed to understand the quirks and twists of
cooperation in scientific communities. It is more likely to say that there is a
desarticulated collection of works, implementing game theory and
complex networks to different situations related to the social life of science.\\

@@

Game theory over complex scientific information and collaboration networks has
also attracted attention, for instance Hanauske has focused on how long-term
strategies may shape different scenarios for Nash equilibria in the interaction
between scientists and journals \cite{Hanauske-etal:2012}. And Hara and
collaborators have used Prisoner's Dilemma in the study of impact factor and
collaboration \cite{Hara-etal:2003,LiebermanNowak:2005}. \\

@@

Even with all these research efforts, cooperation in the context of scientific
collaboration is still loosely defined and the long term dynamics of academic
cooperation (and its consequences) are yet to be fully elucidated. Furthermore,
to our current knowledge, there has been no use of game theory and complex
network analysis for understanding how the topology of scientific collaboration
networks affects access to resources among individuals present in the network
\footnote{For an account of scientific collaboration and definitions, please
  refer to \cite{Sonnenwald:2007}.}. From a different perspective others have
explored the effect of different topologies on the emergence of cooperation
\cite{Santos-etal:2005,HauertDoebeli:2004}. Our work aims to contribute to our
current understanding on the matter, especially when agents have to maximize
their access to resources while taking care of their collaboration links. We
explore the network effect on the distribution of players having access to
certain amount of resources shared by other players in the network and the
distribution of the strength of interactions among them. Particularly, we
implemented two games played simultaneously: one for maximizing individual
utility based on the iterated Prisoner's Dilemma; the other, a coordination game
for maximizing the connection strength between players. We are interested in how
they affect each other in the context of a network of scientific collaboration
under the idea that while researchers are interested in maximizing their
utilities, they also know that it is important to invest in building
collaborative relationships. These two behaviors are explored in a biomedical
research community in M\'exico.\\

>>>>

The analysis of cooperation in scientific research has been the subject of a
number of studies
\cite{VermeulenPenders:2013,Newman:2001,Newman:2004,Elango-etal:2012,HernandezLemus:2013,Strasser:2006,Strasser:2012}. This
is not surprising since cooperation and competition are quite important in
today's academic success. Moreover, the quantitative analysis of cooperation in
science also has a long history of more than 50 years
\cite{Mali-etal:2012,Scharnhorst-etal:2012}. Particularly appealing to
researchers studying this topic is the use of network analysis, especially in
the area of bibliometrics and scientometrics such as in citation networks that
date as back as 1972 \cite{Garfield:1972}, co-authorship and collaboration networks also have a
long tradition \cite{Todorov-etal:1991,Wagner-etal:2017}.\\

The approach we follow here articulates game theory
together with complex network analysis. Putting together game-theoretic modeling
on complex networks is common in studies about the emergence of cooperation
\cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}.
However, in the context of science studies, it is far from being a unified,
active field of research. There is no common set of questions and problems or a
coherent body of models and theory aimed to understand the quirks and twists of
cooperation in scientific communities. It is more likely to say that work of
this kind are scarce and their implementatoin is on different situations related
to the social life of science, where sometimes models are run on science
collaboration networks because access to these kind of data is relatively easy
to access.\\

@@

Game theory over complex scientific information and collaboration networks has
also attracted attention. Hanauske has applied \emph{Evolutionary Game Theory}
to have a better understanding of the market of scientific information, that
involves scientists as producers but also journal, that are in charge of its
dissemination. One of his models connects scientist interest in maximazing their
reputation and their decissions to publish in an Open Access Journal or in a
journal that works under the traditional publication framework. He has also
modeled a game-theretic interactions between scientists and journals for both
players to go open access or not. Only this second model incorporates networks
analysis \cite{Hanauske-etal:2012}.\\

Another interesting behavior explored by game-theoretical modeling and complex
networks has been done by Li and Cao \cite{Li-etal:2009}. This authors analyze the role of
punishment in a science collaboration network. More precisely, they implemented
an ultimatum bargaining model to understand the impact of the topology of a
social network and the emergence of a population of punishers. In the ultimatum
game, there are two players, one of them (the proponent) suggest to split a sum
in two and its connected neighbor (the respondent), can decide to accept or
reject the offer as it may be perceived more or less fair. If the respondent
accepts, the sum is devided and the shares are divided according to the
offer. If the offer is rejected the players do not get anything. According to
the authors, the act of rejecting a selfish offer can be interpreted as a
punishment since the player that rejects it is paying a cost of not receiving
anything but the proponent looses even more, since his offer was selfish. Their
results show that the population of punishers will emerge as a phase transition
as the degree of selfishness increses among players. More over, they found a
dependence relation between the tolerance to selfish behaviors and the
topologies of networks.\\

%%\cite{Hara-etal:2003,LiebermanNowak:2005}. \\

@@

% Even with all these research efforts, cooperation in the context of scientific
% collaboration is still loosely defined and the long term dynamics of academic
% cooperation (and its consequences) are yet to be fully elucidated. Furthermore,
% to our current knowledge, there has been no use of game theory and complex
% network analysis for understanding how the topology of scientific collaboration
% networks affects access to resources among individuals present in the network
% \footnote{For an account of scientific collaboration and definitions, please
%   refer to \cite{Sonnenwald:2007}.}. From a different perspective others have
% explored the effect of different topologies on the emergence of cooperation
% \cite{Santos-etal:2005,HauertDoebeli:2004}.

In a similar vein to the work of Li and Cao, we explore the network effect on
the distribution of players having access to certain amount of resources shared
by other players in the network and the distribution of the strength of
interactions among them. Particularly, we implemented two games played
simultaneously: one for maximizing individual utility based on the iterated
Prisoner's Dilemma; the other, a coordination game for maximizing the connection
strength between players. We are interested in how they affect each other in the
context of a network of scientific collaboration under the idea that while
researchers are interested in maximizing their utilities, they also know that it
is important to invest in building collaborative relationships. These two
behaviors are explored in a biomedical
research community in M\'exico.\\
